---
title: "Multistage Likelihood"
author: "Kính"
output:
    html_document:
        toc: true
        toc_float: true
        code_folding: hide
always_allow_html: true
---

```{r setup, include=F}
here::i_am("code/prep.Rmd")
source('~/Documents/libs.r')
library(forcats)
set.seed(1)
```

## Sample Data

Pull with `rdhs`

```{r adam, eval = F}
library(rdhs)
set_rdhs_config(
    email = "ath19@ic.ac.uk",
    project = "Statistics and Machine Learning for HIV", 
    password_prompt = T
)

ssa <- dhs_countries(returnFields = c("CountryName", "DHS_CountryCode")) %>%
    mutate(regionName = countrycode::countrycode(CountryName, "country.name", "un.regionsub.name")) %>%
    filter(regionName == "Sub-Saharan Africa")

datasets <- dhs_datasets(
    ssa$DHS_CountryCode, 
    surveyYearStart = 2000,
    surveyType = "DHS", 
    fileFormat = "flat", 
    fileType = c("IR", "MR")) %>%
    mutate(sex = if_else(FileType == "Individual Recode", "female", "male"))

downloads <- get_datasets(datasets$FileName)
```


```{r dhs_pull}
fid <- downloads %>% unlist() %>% basename() %>% substr(3, 4) %>% str_detect('IR') %>% which()
flookup <- c(
    weights = "v005", dob = "v011", doi = "v008", lab = "lab", afs = "v531", 
    marriage_age = "v511", n_union = "v503", marital_status = "v501")

o <- map_dfr(fid,
    function(x) {
        readRDS(downloads[[x]]) %>%
            as_tibble() %>%
            mutate(lab = datasets$SurveyId[x]) %>%
            select(any_of(flookup)) %>%
            mutate(
                marital_status = haven::as_factor(marital_status),
                sex = datasets$sex[x]
            )
    }
)
saveRDS(o, here("data/femaleSSA.rds"))

mid <- downloads %>% unlist() %>% basename() %>% substr(3, 4) %>% str_detect('MR') %>% which()
mlookup <- c(weights = "mv005", dob = "mv011", doi = "mv008", lab = 'lab', 
    afs = "mv531", marriage_age = "mv511", n_union = "mv503", marital_status = "mv501")

oo <- map_dfr(mid,
    function(x) {
        readRDS(downloads[[x]]) %>%
            as_tibble() %>%
            mutate(lab = datasets$SurveyId[x]) %>%
            select(any_of(mlookup)) %>% 
            mutate(
                marital_status = haven::as_factor(marital_status),
                sex = datasets$sex[x]
            )
    }
)

saveRDS(oo, here("data/maleSSA.rds"))

# both male and female
ooo <- bind_rows(o, oo)

# recode marital statuses
marital_recode <- c(
   never = "never in union" ,
   never = "never married" ,
   separated = "not living together" ,
   separated = "seperated" ,
   separated = "no longer living together/separated" ,
   union = "living together",
   union = "living with partner",
   NULL = "missing"
)
ooo %<>% mutate(marital_status = fct_recode(marital_status, !!!marital_recode))

#' Data cleaning
#'
ooo %>% 
    mutate(
        n_union = haven::as_factor(n_union), 
        n_union = fct_recode(n_union, NULL = 'missing'),
        age = round((doi - dob)/ 12), 
        afs = if_else(afs > 90, NA_real_, afs), # inconsistent, unknown
        afs = if_else(is.na(afs) & marital_status == "never", 0, afs), # fix missing
        afs = if_else((afs == 0 | is.na(afs)) & marriage_age != 0 & marital_status != "never", marriage_age, afs), # missing afs but aam say otherwise
        marriage_age = if_else(marital_status == "never", 0, marriage_age),
        n_union = if_else(is.na(n_union) & marriage_age == 0, "neverOnce", n_union), 
        n_union = if_else(is.na(n_union) & marriage_age != 0, "oneORmore", n_union), # as censored since missing in data
        across(where(is.labelled), zap_label)
    ) %>%
    filter(!(is.na(marriage_age) & is.na(afs))) %>%  # interim DHS does not have afs & aam, do they named differently
    filter(!is.na(marriage_age)) %>% # 225 had afs but not aam, can be done as censored but just removed here
    saveRDS(here('data/cleaned.rds'))
```
```

Calculate needed indcator for likelihood

```{r likelihood_prep}
readRDS(here('data/cleaned.rds')) %>%
    select(lab, afs, aam = marriage_age, age, u = n_union, ms = marital_status) %>% 
    mutate(
        across(where(is.labelled), zap_labels),
        cc = substr(lab, 1, 2),
        ms = if_else(ms == "union", "married", ms),
        afs = if_else(afs != 0 & aam != 0 & afs >= aam, aam, afs), # remove VX if VM
        VV = if_else(afs != 0, afs - 1, age - 1),
        VX = if_else(afs != 0 & afs != aam, afs, -2908),
        VM = if_else(afs != 0 & aam != 0 & afs == aam, aam, -2908),
        XM = if_else(afs != 0 & aam != 0 & afs != aam, aam, -2908),
        XX = case_when(
            afs != 0 & aam != 0 & afs != aam ~ aam - afs, 
            afs != 0 & aam == 0 & afs != age ~ age - afs, 
            otherwise ~ -2908),
        MM = if_else(aam != 0 &   u == "once" & ms == "married", age - aam, -2908),
        MJ = if_else(aam != 0 & !(u == "once" & ms == "married"), aam, -2908),
        delta = if_else(MJ > 0, age - aam + 1, -2908), 
        J = case_when(
            delta > 0 & ms == "separated" ~ 4,
            delta > 0 & ms == "divorced" ~ 5,
            delta > 0 & ms == "widowed" ~ 6,
            delta > 0 & ms == "married" ~ 7,
            otherwise ~ -2908)
    ) %>% 
    count(cc, age, afs, aam, VV, VX, XX, XM, VM, MM, MJ, J, delta) %>% 
    rename(count = n) %T>% # note side effect here
    allot(tmp) %>% 
    saveRDS(here::here('data/d_agg.rds'))

# add tests later if found
mustbe <- function(x, y) expect_that(x, testthat::equals(y))
tmp %>% filter(afs == 0, XX > 0) %>% nrow() %>% mustbe(0)
tmp %>% filter(XM > 0, afs == 0) %>% nrow() %>% mustbe(0)

```